---
title: "Introduction to easytable"
author: "Alfredo Hernandez Sanchez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to easytable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

The **easytable** package creates publication-ready regression tables in multiple formats from a single function call. This vignette demonstrates the key features using the Palmer Penguins dataset.

## Installation

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("alhdzsz/easytable")
```

## Setup

```{r message=FALSE}
library(easytable)
library(palmerpenguins)
data(penguins)
```

## Basic Usage

Let's start with a simple example using the Palmer Penguins dataset. We'll model penguin body mass as a function of flipper length, species, and island.

```{r}
# Fit three nested models
m1 <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
m2 <- lm(body_mass_g ~ flipper_length_mm + species, data = penguins)
m3 <- lm(body_mass_g ~ flipper_length_mm + species + island, data = penguins)

# Create named list
models <- list(
  "Model 1" = m1,
  "Model 2" = m2,
  "Model 3" = m3
)
```

### Word Output (Default)

The default output format is Word, using the flextable package:

```{r eval=FALSE}
easy_table(models)
```

This creates a formatted table that can be:
- Viewed in RStudio
- Saved to Word with `flextable::save_as_docx()`
- Included in R Markdown Word documents

### Markdown Output

For Quarto and R Markdown HTML/PDF documents:

```{r}
easy_table(models, output = "markdown")
```

### LaTeX Output

For academic papers and LaTeX documents:

```{r}
easy_table(models, output = "latex")
```

## Robust Standard Errors

Use heteroskedasticity-consistent standard errors with the `robust.se` parameter:

```{r eval=FALSE}
easy_table(models, output = "markdown", robust.se = TRUE)
```

This automatically uses HC-robust standard errors via the `lmtest` and `sandwich` packages. The output includes a note indicating robust SEs were used.

**When to use robust SE:**
- Heteroskedasticity is suspected or detected
- Large sample sizes where asymptotic properties apply
- Published standards in your field require it

## Marginal Effects

Instead of raw coefficients, you can report average marginal effects (AME):

```{r eval=FALSE}
# Marginal effects (note: requires margins package)
easy_table(models[2:3], output = "markdown", margins = TRUE)
```

**Important notes about marginal effects:**
- Intercept is excluded (not meaningful for AME)
- Use Model 2+ (need at least one covariate beyond constant)
- Interpretation: unit change in X → change in E[Y]

## Combining Robust SE and Margins

You can use both together:

```{r eval=FALSE}
easy_table(
  models[2:3],
  output = "markdown",
  robust.se = TRUE,
  margins = TRUE
)
```

This computes marginal effects with robust standard errors.

## Control Variables

When you have many control variables (e.g., many factor levels), you can group them to show presence/absence rather than individual coefficients:

```{r}
# Add more controls
m4 <- lm(body_mass_g ~ flipper_length_mm + bill_length_mm, data = penguins)
m5 <- lm(body_mass_g ~ flipper_length_mm + bill_length_mm + species + island,
         data = penguins)

models_controls <- list(
  "Baseline" = m4,
  "With Controls" = m5
)

# Group species and island as control variables
easy_table(
  models_controls,
  output = "markdown",
  control.var = c("species", "island")
)
```

Instead of showing:
- `speciesChinstrap`
- `speciesGentoo`
- `islandDream`
- `islandTorgersen`

The table shows:
- `species: Y` (indicating presence)
- `island: Y` (indicating presence)

This is especially useful with:
- Factor variables with many levels
- Time or location fixed effects
- Any variable where you want to indicate presence without showing all coefficients

## Highlighting Significant Results

You can color-code significant results (works best with Word output):

```{r eval=FALSE}
easy_table(models, output = "word", highlight = TRUE)
```

- **Green background**: Positive significant coefficients
- **Red background**: Negative significant coefficients

## GLM Models

The package also supports generalized linear models:

```{r}
# Logistic regression for sex (remove NAs first)
penguins_complete <- na.omit(penguins)

g1 <- glm(sex ~ body_mass_g, data = penguins_complete, family = binomial)
g2 <- glm(sex ~ body_mass_g + species, data = penguins_complete, family = binomial)

glm_models <- list(
  "Bivariate" = g1,
  "With Species" = g2
)

easy_table(glm_models, output = "markdown")
```

## Exporting to CSV

You can export the underlying data table to CSV:

```{r eval=FALSE}
easy_table(
  models,
  output = "word",
  csv = "penguin_regression_results"
)
# Creates: penguin_regression_results.csv
```

This is useful for:
- Archiving raw results
- Further processing in other software
- Sharing with collaborators who need raw numbers

## Complete Example

Here's a comprehensive example combining all features:

```{r eval=FALSE}
# Fit models with controls
m_baseline <- lm(body_mass_g ~ flipper_length_mm + bill_length_mm,
                 data = penguins)

m_full <- lm(body_mass_g ~ flipper_length_mm + bill_length_mm +
               species + island + sex,
             data = na.omit(penguins))

models_full <- list(
  "Baseline" = m_baseline,
  "Full Model" = m_full
)

# Create publication-ready table
easy_table(
  models_full,
  output = "word",
  robust.se = TRUE,
  control.var = c("species", "island", "sex"),
  highlight = TRUE,
  csv = "final_results"
)
```

## Interpretation Guide

### Reading the Output

Each cell shows:
```
Estimate ***
(Std. Error)
```

Where:
- `***` = p < 0.01
- `**` = p < 0.05
- `*` = p < 0.1
- (blank) = p > 0.1

### Model Fit Statistics

At the bottom of each table:
- **N**: Number of observations
- **R sq.**: R-squared (proportion of variance explained)
- **Adj. R sq.**: Adjusted R-squared (penalized for number of predictors)
- **AIC**: Akaike Information Criterion (lower is better; for model comparison)

## Comparison of Output Formats

| Format | Best For | Pros | Cons |
|--------|----------|------|------|
| **Word** | MS Word docs, presentations | Beautiful formatting, easy to customize | Requires flextable |
| **Markdown** | Quarto/RMarkdown reports | Plain text, version control friendly | Basic formatting |
| **LaTeX** | Academic papers, journals | Professional typesetting, journal standards | Requires LaTeX knowledge |

## Tips and Best Practices

### Model Naming

Use descriptive model names:

```{r}
# Good
models_named <- list(
  "Baseline" = m1,
  "Add Species" = m2,
  "Full Model" = m3
)

# Less clear
models_numbered <- list(
  "Model1" = m1,
  "Model2" = m2,
  "Model3" = m3
)
```

### When to Use Control Variables

Use `control.var` when:
- You have many factor levels (e.g., 50 states, 100 countries)
- Including fixed effects (time, location)
- Control variables aren't your main focus
- You want a more compact table

Don't use `control.var` when:
- You want to show specific coefficient estimates
- You have only 2-3 levels
- The variable is a key predictor of interest

### Robust Standard Errors

In general, use `robust.se = TRUE` when:
- You have cross-sectional data with potential heteroskedasticity
- Sample size is reasonably large (n > 30)
- You want to be conservative

Don't use when:
- You've explicitly modeled heteroskedasticity (e.g., WLS)
- Small sample sizes (asymptotic properties may not hold)
- Time series data (use HAC errors instead, not currently supported)

## Troubleshooting

### Missing Package Errors

If you get an error like:
```
Package 'flextable' is required for Word output.
  Install it with: install.packages('flextable')
```

Simply install the suggested package:
```{r eval=FALSE}
install.packages("flextable")
```

### Model Compatibility

Currently supported:
- `lm()` - linear regression
- `glm()` - generalized linear models

Coming soon: `plm`, `fixest`, and other model types.

## Next Steps

- See `?easy_table` for complete function documentation
- Check [GitHub](https://github.com/alhdzsz/easytable) for latest updates
- Report issues or request features on GitHub Issues

## References

- Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0.
- Leeper TJ (2021). margins: Marginal Effects for Model Objects. R package.
- Zeileis A (2004). "Econometric Computing with HC and HAC Covariance Matrix Estimators." Journal of Statistical Software, 11(10), 1-17.
