---
title: "Penguins Tutorial: Easy to Use, Easy to Read"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Penguins Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This tutorial shows the full `easytable` workflow using `palmerpenguins`.

## 1) Fit a sequence of models

```{r}
library(easytable)
library(palmerpenguins)

m1 <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
m2 <- lm(body_mass_g ~ flipper_length_mm + species, data = penguins)
m3 <- lm(body_mass_g ~ flipper_length_mm + species + island, data = penguins)
```

## 2) Create a readable baseline table

```{r}
easytable(
  m1, m2, m3,
  model.names = c("Baseline", "With Species", "Full Model")
)
```

By default, each coefficient cell is two lines:

1. Estimate + significance stars
2. Standard error in parentheses

## 3) Highlight significant coefficients

```{r}
easytable(
  m1, m2, m3,
  model.names = c("Baseline", "With Species", "Full Model"),
  highlight = TRUE
)
```

## 4) Collapse controls into indicator rows

```{r}
easytable(
  m1, m2, m3,
  model.names = c("Baseline", "With Species", "Full Model"),
  control.var = "island",
  highlight = TRUE
)
```

`control.var` is useful when models include many factor levels or fixed effects.

## 5) LaTeX output for PDF workflows

```{r}
easytable(
  m1, m2, m3,
  model.names = c("Baseline", "With Species", "Full Model"),
  output = "latex",
  control.var = "island",
  highlight = TRUE
)
```

## 6) Robust standard errors (optional dependency path)

```{r}
if (requireNamespace("lmtest", quietly = TRUE) &&
    requireNamespace("sandwich", quietly = TRUE)) {
  easytable(
    m1, m2, m3,
    model.names = c("Baseline", "With Species", "Full Model"),
    robust.se = TRUE
  )
}
```

## 7) Marginal effects (optional dependency path)

```{r}
if (requireNamespace("margins", quietly = TRUE)) {
  easytable(
    m2, m3,
    model.names = c("With Species", "Full Model"),
    margins = TRUE
  )
}
```

## 8) Export outputs

```{r eval = FALSE}
easytable(
  m1, m2, m3,
  model.names = c("Baseline", "With Species", "Full Model"),
  highlight = TRUE,
  export.word = "penguins_table.docx",
  export.csv = "penguins_table.csv"
)
```

## Design Notes

`easytable` keeps these display invariants:

1. Two-line coefficient cells.
2. Zebra only in coefficient rows.
3. One divider between coefficient rows and model-stat rows.
4. Control indicators in the model-stat block.

These defaults are intentional so tables remain legible in long workflows.
